FTLR proof cases

### 2018-06-24 fully-abstract-comp/ftlr-proof-cases.tex
The last of the FTLR call case.

Case: \ref{case:ftlr:scall:mem}, we need to show

\begin{equation}
  \memSat[n'-1]{\Phi_S''.\mem,\Phi_S''.\stk,\Phi_S''.\ms_\stk,\Phi_T''.\mem}{W_M'}\label{eq:ftlr:call:memsat-case}
\end{equation}
We have the following memory satisfaction on the initial memory:
\begin{equation}
  \memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}\label{eq:ftlr-call-og-mem-sat}
\end{equation}
By definition of memory satisfaction, this gives us:
\begin{itemize}
\item $\stk = (\opc_1,\ms_1) :: \dots :: (\opc_m,\ms_m)$
\item $\ms_1 \# \dots \# \ms_m \# \ms_S \# \ms_\stk$
\item $W = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}}$
\item $\ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}, \ms_{T,f}, \ms_{S,f}, \ms_S',\overline{\sigma}$ such that
  \begin{itemize}
  \item $\ms_S = \ms_{f,S} \uplus \ms_S'$
  \item $\ms_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus \ms_\var{T,heap} \uplus \ms_{T,f}$
  \item $\memSatStack{\stk,\ms_\var{T,stack}}{W_{\var{stack}}}$
  \item $\memSatFStack{\ms_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}}$
  \item $\npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(\pwheap)(W_{\var{heap}})$
  \end{itemize}
\end{itemize}
We need to consider the memories of $\Phi_S''$ and $\Phi_T''$.
The memories of the initial configurations have only changed little:
\[
  \Phi_T''.\mem = \Phi_T''.\mem\update{\aaddr_\stk}{42}
\]
and 
\[
\Phi_S''.\ms_\stk = \Phi_S.\ms_\stk - \Phi_S.\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]}
\]
and
\[
  \Phi_S''.\stk = ((\aaddr+\calllen),\ms_{\stk\_\priv,S}) :: \Phi_S.\stk
\]
with $\ms_{\stk\_\priv,S} = \Phi_S.\ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42}$.
We need to show the memory satisfaction in \ref{eq:ftlr:call:memsat-case}. Say
$(\opc_0,\ms_0) = ((\aaddr+\calllen),\ms_{\stk\_\priv,S})$. In this case we have
to argue 
\[
  \ms_{\stk\_\priv,S} \# \ms_1 \# \dots \# \ms_m \# \ms_S \# \Phi_S.\ms_\stk - \Phi_S.\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]}
\]
which follows from Assumption~\ref{eq:ftlr:call:memsat-case} and the fact that the memory we remove from $\Phi_S.\ms_\stk$ is exactly what we add to the private stack.
\lau{This is a bit more involved than I make it to be here, but the fact that we also have a safety assumption on the register-file with the stack pointer means that all of the memory removed from the free stack is actually there.}

  This concludes case \ref{case:ftlr:scall:reg}

  Case: \ref{case:ftlr:scall:mem}, \lau{TODO}
\end{proof}

@@

\begin{lemma}
  \label{lem:stack-sat-unchanged-future}
  If $\pwfree = \pwfree[W']$ and $W' \future W$ and
  $\memSatFStack{\ms_\stk,\ms_T}{W}$, then $\memSatFStack{\ms_\stk,\ms_T}{W'}$.
\end{lemma}

>>>>

% Case: \ref{case:ftlr:scall:mem}, we need to show

% \begin{equation}
%   \memSat[n-1]{\Phi_S''.\mem,\Phi_S''.\stk,\Phi_S''.\ms_\stk,\Phi_T''.\mem}{W_M'}\label{eq:ftlr:call:memsat-case}
% \end{equation}
% We have the following memory satisfaction on the initial memory:
% \begin{equation}
%   \memSat[n]{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}\label{eq:ftlr-call-og-mem-sat}
% \end{equation}
% By definition of memory satisfaction, this gives us:
% \begin{itemize}
% \item $\stk = (\opc_1,\ms_1) :: \dots :: (\opc_m,\ms_m)$
% \item $\ms_1 \# \dots \# \ms_m \# \ms_S \# \ms_\stk$
% \item $W = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}}$
% \item $\ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}, \ms_{T,f}, \ms_{S,f}, \ms_S',\overline{\sigma}$ such that
%   \begin{itemize}
%   \item $\ms_S = \ms_{f,S} \uplus \ms_S'$
%   \item $\ms_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus \ms_\var{T,heap} \uplus \ms_{T,f}$
%   \item $\memSatStack{\stk,\ms_\var{T,stack}}{W_{\var{stack}}}$
%   \item $\memSatFStack{\ms_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}}$
%   \item $\npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(\pwheap)(W_{\var{heap}})$
%   \end{itemize}
% \end{itemize}
% We need to consider the memories of $\Phi_S''$ and $\Phi_T''$.
% The memories of the initial configurations have only changed little:
% \[
%   \Phi_T''.\mem = \Phi_T''.\mem\update{\aaddr_\stk}{42}
% \]
% and 
% \[
% \Phi_S''.\ms_\stk = \Phi_S.\ms_\stk - \Phi_S.\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]}
% \]
% and
% \[
%   \Phi_S''.\stk = ((\aaddr+\calllen),\ms_{\stk\_\priv,S}) :: \Phi_S.\stk
% \]
% with $\ms_{\stk\_\priv,S} = \Phi_S.\ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42}$.
% We need to show the memory satisfaction in \ref{eq:ftlr:call:memsat-case}. Say
% $(\opc_0,\ms_0) = ((\aaddr+\calllen),\ms_{\stk\_\priv,S})$. In this case we have
% to argue 
% \[
%   \ms_{\stk\_\priv,S} \# \ms_1 \# \dots \# \ms_m \# \ms_S \# \Phi_S.\ms_\stk - \Phi_S.\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]}
% \]
% which follows from Assumption~\ref{eq:ftlr:call:memsat-case} and the fact that the memory we remove from $\Phi_S.\ms_\stk$ is exactly what we add to the private stack.
% \lau{This is a bit more involved than I make it to be here, but the fact that we also have a safety assumption on the register-file with the stack pointer means that all of the memory removed from the free stack is actually there.}

%   This concludes case \ref{case:ftlr:scall:reg}

  Case \ref{case:ftlr:scall:mem}:
  we need to show:
  \[
    \memSat[n-1]{\Phi_S''.\mem,\Phi_S''.\stk,\Phi_S''.\ms_\stk,\Phi_T''.\mem}{W_M'}
  \]
  which amounts to
  \[
    \memSat[n-1]{\Phi_S.\mem,((\aaddr+\calllen),\ms_{\stk\_\priv,S}) :: \Phi_S,\Phi_S.\ms_\stk - \Phi_S.\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]},\Phi_T''.\mem}{W_M'}
  \]
  for $\ms_{\stk\_\priv,S} = \Phi_S.\ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42}$.

  In order to show this, we will first show the following:
  \begin{itemize}
  \item $\baddr_\stk = \stkb$

    We know $\Phi_S$ is reasonable up to $n$ steps.
    Further, from $\callCond{\Phi_S,r_1,r_2,a}$ and $b \leq a$ and $a + \calllen -1 \leq e$ and $[b,e] \subseteq \ta$ we can conclude that $\Phi_S$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ in $\ta$.
    By the \emph{Guarantee stack base address before call} we then know $\Phi_S(r_\stk) = \stkptr{\_,\stkb,\_,\_}$.
  \end{itemize}

  By assumption we have
  $\memSat{\Phi_S.\mem,\Phi_S.\stk,\Phi_S.\ms_\stk,\Phi_T.\mem}{W_M}$ which
  gives us $\ms_\var{T,stack}, \ms_\var{T,free\_stack},$ $\ms_\var{T,heap},
  \ms_{T,f}, \ms_{S,f}, \ms_S', W_{M,\var{stack}}, W_{M,\var{free\_stack}},
  W_{M,\var{heap}}$ such that:
  \begin{itemize}
  \item $\Phi_S.\stk = (\opc_0, \ms_0) \dots (\opc_m,\ms_m)$
  \item $\ms_0 \uplus \dots \uplus \ms_m \uplus \Phi_S.\mem \uplus \Phi_S.\ms_\stk$
  \item $W_M = W_{M,\var{stack}} \oplus W_{M,\var{free\_stack}} \oplus W_{M,\var{heap}}$
  \item $\Phi_S.\mem = \ms_{f,S} \uplus \ms_S'$
  \item $\Phi_T.\mem = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus
    \ms_\var{T,heap} \uplus \ms_{T,f}$
  \item $\memSatStack{\Phi_S.\stk,\ms_\var{T,stack}}{W_{M,\var{stack}}}$
  \item $\memSatFStack{\Phi_S.\ms_\stk,\ms_\var{T,free\_stack}}{W_{M,\var{free\_stack}}}$
  \item $\npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(\pwheap)(W_{M,\var{heap}})$
  \end{itemize}

  In order to prove the memory satisfaction, we pick the same memories except for the following changes:
  \begin{itemize}
  \item The target free stack partition: $\ms_\var{T,free\_stack} \setminus \ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]}$
  \item The target private stack partition: $\ms_\var{T,stack} \uplus \ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42}$
  \end{itemize}
  and the worlds
  \begin{itemize}
  \item Free stack world: $W_{M,\var{free\_stack}}'$ is $W_{R,1}'$ with the ownership of $W_{M,\var{free\_stack}}$ except that it gives up any ownership that we used for the safety of the addresses $[\aaddr_\stk,\eaddr_\stk]$.
  \item Private stack world: $W_{M,\var{stack}}'$ is $W_{R,1}'$ with the ownership of $W_{M,\var{stack}}$ except that it takes the ownership that $W_{M,\var{free\_stack}}$ used for the safety of the addresses $[\aaddr_\stk,\eaddr_\stk]$.
  \item Heap world: $W_{M,\var{heap}}'$ is $W_{R,1}'$ with the ownership of $W_{M,\var{heap}}$ which gives us $W_{M,\var{heap}}' \future W_{M,\var{heap}}$.
  \end{itemize}
  We now need to show:
  \begin{itemize}
  \item $\stk = ((\aaddr+\calllen),\ms_{\stk\_\priv,S}) :: (\opc_0, \ms_0) \dots (\opc_m,\ms_m)$

    Trivial.
  \item $\ms_{\stk\_\priv,S} \uplus \ms_0 \uplus \dots \uplus \ms_m \uplus \Phi_S.\mem \uplus \Phi_S.\ms_\stk - \Phi_S.\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]}$

    Follows by assumption and the fact that we have shuffled around some memory\footnote{Thanks to the register-file safety assumption we know for sure that the memory we remove from the free stack is actually there.}.
  \item $W_M' = W_{M,\var{stack}}' \oplus W_{M,\var{free\_stack}}' \oplus W_{M,\var{heap}}'$

    This follows by assumption and the fact that all ownership we have added to a world has been removed from another.
  \item $\Phi_S.\mem = \ms_{f,S} \uplus \ms_S'$

    By assumption.
  \item $\Phi_T.\mem = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42} \uplus \ms_\var{T,free\_stack} \setminus \ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]} \uplus \ms_\var{T,heap} \uplus \ms_{T,f}$

    Follows by assumption and the fact that we have shuffled around some memory\footnote{Thanks to the register-file safety assumption we know for sure that the memory we remove from the free stack is actually there.}.
  \item $\memSatStack{((\aaddr+\calllen),\ms_{\stk\_\priv,S}) :: \Phi_S,\ms_\var{T,stack} \uplus \ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42}}{W_{M,\var{stack}}'}$:

    For most of the conditions, they follow from the previous stack satisfaction assumption.
    The only challenge is to argue that the new stack frame satisfies the conditions.
    First of all, we know $\ms_{\stk\_\priv,S}$ is non-empty as it at least contains address $\aaddr_\stk$.
    Next, we need to argue that
    \[
      \begin{array}{l}
        \forall i \in \{0,\dots,m\} \ldotp \\
        \quad \forall a \in \dom(\ms_{\stk\_\priv,S}) \ldotp \forall a' \in \dom(\ms_j) \ldotp \stkb < a < a') \wedge\\
      \end{array}
    \]
    The first bit, $\stkb < a$ for $a \in \dom(\ms_{\stk\_\priv,S})$ follows from the fact that $a_\stk$ is the smallest address of $a \in \dom(\ms_{\stk\_\priv,S})$ and by assumption $\stkb < a_\stk$.

    Now assume for contradiction that there exists $a' \in \dom(\ms_j)$ for some $j$ such that $a' \le a$ for some $a \in \dom(\ms_{\stk\_\priv,S})$.
    By assumption we have $\stkb < a'$ so $a' \in [\stkb+1,a] \subseteq [\stkb+1,a_\stk] \subseteq [\stkb,e_\stk]$ which means that $a'$ is an address governed by a stack pointer.
    By the register-safety assumption this must mean that it is an address of the free part of the stack.
    At the same time, it must be an address of the private stack because $\ms_j$ is part of the stack from the original configuration.
    This contradicts the initial memory satisfaction assumption as the different parts must be disjointed.

    From the stack satisfaction assumption, we get $R_\ms$ and $R_W$. Pick
    \[
      R_\ms'(r) =
      \begin{cases}
        (\ms_{\stk\_\priv,S}, (\aaddr+\calllen),\ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42})& \text{for $r = r_\var{priv\_stk}$}\\
        R_\ms(r) & \totherwise
      \end{cases}
    \]
    and for $R_W'$ pick
    \[
      R_W'(r) =
      \begin{cases}
        W_{M,\var{free\_stack},[\aaddr_\stk,\eaddr_\stk]}[\free.R^{-1}([\aaddr_\stk,\eaddr_\stk]) \mapsto \revoked][\priv.r_\var{priv\_stk} \mapsto (\stareg[(\ms_{\stk\_\priv,S},\Phi_T''.\mem|_{[\aaddr_\stk,\eaddr_\stk]}),\gc]{\spa,\square},\aaddr+\calllen)]
        & \text{for $r = r_\var{priv\_stk}$} \\
        R_W(r)[\free.R^{-1}([\aaddr_\stk,\eaddr_\stk]) \mapsto \revoked][\priv.r_\var{priv\_stk} \mapsto (\stareg[(\ms_{\stk\_\priv,S},\Phi_T''.\mem|_{[\aaddr_\stk,\eaddr_\stk]}),\gc]{\spa,\square},\aaddr+\calllen)] &
      \end{cases}
    \]
    where $W_{M,\var{free\_stack},[\aaddr_\stk,\eaddr_\stk]}$ is the world that the free stack assumption uses to satisfy that range of addresses.

    Most of the condition trivially holds. 
    % This follows in part by mono of the interpretation function and the worlds
    % picked are future worlds of the past worlds.
    The only one that requires some argumentation is
    \[
      \npair[n-1]{(\ms_{\stk\_\priv,S},\ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42})}
      \in
      \stareg[(\ms_{\stk\_\priv,S},\Phi_T''.\mem|_{[\aaddr_\stk,\eaddr_\stk]}),\gc]{\spa,\square}.H
      \; \xi^{-1}(R_W'(r_\var{priv\_stk}))
    \]
    Where $R_W'(r_\var{priv\_stk})$ is the part of $W_{M,\var{stack}}'$ with the ownership used for addresses $[\aaddr_\stk, \eaddr_\stk]$ in the memory satisfaction assumption.

    As we have the safe register-file assumption, we know that the stack capability is safe.
    This means that addresses $[\aaddr_\stk,\eaddr_\stk]$ must be part of the free stack.
    Further, $\Phi''_T.\mem = \Phi_T.\mem\update{\aaddr_\stk}{42}$ and $\ms_\var{T,free\_stack} \subseteq \Phi_T.\mem$ and $\dom(\ms_\var{T,free\_stack}) \supseteq [\aaddr_\stk,\eaddr_\stk]$ from which it follows that the memories are equal to the one of the static region.

    It remains to show
    \[
      \forall \aaddr \in [\aaddr_\stk,\eaddr_\stk] \ldotp \npair[n-1]{(\ms_{\stk\_\priv,S}(\aaddr),\ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42}(\aaddr))} \in \lrv(W_{R,\aaddr}')
    \]
    for $\aaddr = \aaddr_\stk$ it is trivial as we have to show
    \[
      \npair[n-1]{(42,42)} \in \lrv(W_{R,\aaddr_\stk}')
    \]
    for $\aaddr \in [\aaddr_\stk+1,\eaddr_\stk]$ we need to show
    \[
      \npair[n-1]{(\ms_{\stk\_\priv,S}(\aaddr),\ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42}(\aaddr))} \in \lrv(W_{R,\aaddr}')
    \]
    If we can show
    \[
      \npair[n-1]{(\Phi_S.\ms_\stk(\aaddr),\ms_\var{T,free\_stack}(\aaddr))} \in \lrv(W_{R,\aaddr})
    \]
    then we are done by monotonicity of $\lrv$.

    By assumption we know $\npair{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$ which entails $\npair{(\Phi_S.\reg(\rstk),\Phi_T.\reg(\rstk))} \in \lrrg{\trust}(W_{R,\rstk})$.
    We know $\Phi_S.\reg(\rstk) = \stkptr{\rw,\stkb,\eaddr_\stk,\aaddr_\stk}$, so by the definition of $\lrv$ we get
    \[
      \npair{[\stkb,\eaddr_\stk]} \in \stackReadCond{W_{R,\rstk}}
    \]
    which in turn gives us $S_{\rstk} \subseteq \addressable{\linear,\pwfree}$ and
    $R_{\rstk} : S_{\rstk} \fun \powerset{\nats}$ for which
    \begin{itemize}
    \item $\forall r \in S_{\rstk} \ldotp |R_{\rstk}(r)| = 1$
    \item $\uplus_{r \in S_{\rstk}} R_{\rstk}(r) \supseteq [\stkb,\eaddr_\stk]$
    \item $\forall r \in S_{\rstk} \ldotp W_{R,\rstk}(r).H \nsubeq \stdreg{R_{\rstk}(r),\gc}{\spao}$
    \end{itemize}
    Further, we know $\memSatFStack{\Phi_S.\ms_\stk,\ms_\var{T,free\_stack}}{W_{M,\var{free\_stack}}.\free}$ which means that we have $R_\ms : \dom(\activeReg{W_{M,\var{free\_stack}}}) \fun MemSeg \times \MemSeg$ and $R_W : \dom(W_{M,\var{free\_stack}}.\free) \fun \Worlds$.
    $R_W$ distributes the ownership of $W_{M,\var{free\_stack}}$ and $R_\ms$ partitions the memories.

    We know that all regions in $S_{\rstk}$ govern singleton memory segments, so $R_\ms$ must map to singleton memory segment pairs for $r \in S_{\rstk}$.
    Further, by definition of the free stack satisfaction for $r \in \S_{\rstk}$ we have
    \[
      \npair{R_\ms(r)} \in W_{R,\rstk}(r).H \; \xi^{-1}(R_W(r))
    \]
    which entails
    \[
      \npair[n-1]{R_\ms(r)} \in \stdreg{R_{\rstk}(r),\gc}{\spao} \; \xi^{-1}(R_W(r))
    \]
    which entails
    \[
      \npair[n-1]{R_\ms(r)(a)} \in \lrv (R_W(r))
    \]
    for $a \in R_{\rstk(r)}$.

    Now this is exactly what we wanted to show because $R_W(r)$ is what we picked as $W_{R,\aaddr}$.
  \item $\memSatFStack{\Phi_S.\ms_\stk - \Phi_S.\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]},\ms_\var{T,free\_stack} \setminus \ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]}}{W_{M,\var{free\_stack}}'}$:

    From the safety assumption on the stack capability, we can deduce a number of things:
    \begin{itemize}
    \item $[\aaddr_\stk,\eaddr_\stk]$ must have been part of the free stack
    \item for every address in $[\aaddr_\stk,\eaddr_\stk]$ there is a region for that singleton memory segment.
    \end{itemize}
    The first part means that we do indeed remove all of the memory we try to subtract.
    The latter means that we can reuse the same split of the remaining memory and the world ownership as we get from assumption $\memSatFStack{\Phi_S.\ms_\stk,\ms_\var{T,free\_stack}}{W_{M,\var{free\_stack}}}$.
    Using this, the result follows from monotonicity of the $H$ function.

  \item $\npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(\pwheap)(W_{M,\var{heap}}')$:

    Follows from Lemma~\ref{lem:heap-sat-unchanged-future}, the fact the the heap part of the world remains unchanged and $W_{M,\var{heap}}' \future W_{M,\var{heap}}$.
  \end{itemize}
\end{proof}

@@

\begin{lemma}
  \label{lem:stack-sat-unchanged-future}
  If $\pwfree = \pwfree[W']$ and $W' \future W$ and
  $\memSatFStack{\ms_\stk,\ms_T}{W}$, then $\memSatFStack{\ms_\stk,\ms_T}{W'}$.
\end{lemma}
\begin{proof}
  TODO
\end{proof}

\begin{lemma}
  \label{lem:heap-sat-unchanged-future}
  If $\pwheap = \pwheap[W']$ and $W' \future W$ and $\npair{(\overline{\sigma},\ms_S,\ms_T)} \in \lrheap(W)$, then $\npair{(\overline{\sigma},\ms_S,\ms_T)} \in\lrheap(W')$.
\end{lemma}
\begin{proof}
  TODO
\end{proof}
